Numerical simulations of the Cordilleran ice sheet through the last glacial cycle

### 2015-05-28 cordillera-cycle.tex
Add calving section.

Ice sheet dynamics are typically modelled using a combination of internal
deformation and basal sliding. PISM is a~shallow ice sheet model, which implies
that the balance of stresses is approximated based on their predominant
components. The Shallow Shelf Approximation (SSA) is used as a ``sliding law''
for the Shallow Ice Approximation (SIA) by adding velocity solutions of the
two approximation
\citep[Eqns.~7--9 and 15]{Bueler.Brown.2009, Winkelmann.etal.2011}.
Ice rheology depends on temperature and water content through an enthalpy
formulation \citep{Aschwanden.etal.2012}. Surface air temperature derived from
the climate forcing (Sect.~\ref{sec:atm}) provides the upper boundary
condition to the ice enthalpy model. Temperature is computed subglacially to
a~depth of 3\,km,
where it is conditioned by a lower boundary geothermal heat flux of
70\,\unit{mW\,m^{-2}}. Although this uniform value does not account for the
high spatial geothermal variability in the region
\citep{Blackwell.Richards.2004}, it is, on average, representative of available
heat flow measurements. In the low-resolution simulations, the vertical grid
consists of up to 51~enthalpy layers in the ice sheet and 31~temperature layers
in the bedrock. In the high-resolution simulations, up to 101~ice layers, and
61~bedrock layers are used.

@@

>>>>

Ice sheet dynamics are typically modelled using a combination of internal
deformation and basal sliding. PISM is a~shallow ice sheet model, which implies
that the balance of stresses is approximated based on their predominant
components. The Shallow Shelf Approximation (SSA) is used as a ``sliding law''
for the Shallow Ice Approximation (SIA) by adding velocity solutions of the
two approximation
\citep[Eqns.~7--9 and 15]{Bueler.Brown.2009, Winkelmann.etal.2011}.

Ice rheology depends on temperature and water content through an enthalpy
formulation \citep{Aschwanden.etal.2012}. Surface air temperature derived from
the climate forcing (Sect.~\ref{sec:atm}) provides the upper boundary
condition to the ice enthalpy model. Temperature is computed subglacially to
a~depth of 3\,km,
where it is conditioned by a lower boundary geothermal heat flux of
70\,\unit{mW\,m^{-2}}. Although this uniform value does not account for the
high spatial geothermal variability in the region
\citep{Blackwell.Richards.2004}, it is, on average, representative of available
heat flow measurements. In the low-resolution simulations, the vertical grid
consists of up to 51~enthalpy layers in the ice sheet and 31~temperature layers
in the bedrock. In the high-resolution simulations, up to 101~ice layers, and
61~bedrock layers are used.

@@

Ice-shelve calving is computed by a double criterion. First, a
physically-realistic calving flux is computed based on eigenvalues of the
horizontal strain rate tensor \citep{Winkelmann.etal.2011,
Levermann.etal.2012}. This allows outlet glaciers to advance in confined
embayments, but prevents extensive ice-shelves to build-up in the open ocean.
Second, floating ice thinner than 50\,m is systematically calved off. A subgrid
scheme \citep{Albrecht.etal.2011} allows for a continuous migration of the
calving frontm. Such treatment of calving was applied to the Antarctic ice
sheet and shown to produce realistic calving front positons for many of the
ice-shelves \citep{Martin.etal.2011}.
